---
title: "HW02"
author: Noah Cowit
date: "September 17 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1
##a. 
If a and c have different signs, the correlation between V amd W will be inverted.
$$\begin{aligned}
Cor(V,W)&=\frac{Cov(aX+b,cY+d)}{SD(aX+b)SD(cY+d)}\\
&=\frac{acCov(XY)}{ac\sigma_{X}\sigma_{Y}}\\
&=\frac{Cov(XY)}{\sigma_{X}\sigma_{Y}}\\
&=\pm Cov(X,Y)
\end{aligned}$$
   
###
##b.
$$\begin{aligned}
r_{VW}=\frac{\sum z_{V_i}z_{W_i}}{n-1}&=\frac{\sum \frac{((aX+b)_i-a\overline X-b)}{\sigma_{aX}}\frac{((cY+d)_i-c\overline Y-d)}{\sigma_{cY}}} {n-1}\\
&=\frac{\pm \sum \frac{((X)_i-\overline X)}{\sigma_{X}}\frac{((Y)_i\overline Y)}{\sigma_{Y}}} {n-1}\\
&=\pm r_{XY}
\end{aligned}$$
   
###
##c.
$$\begin{aligned}
b0_{XY}&=r_{XY}\frac{\sigma_{Y}}{\sigma_{X}}\\
b1_{WV}&=r_{WV}\frac{\sigma_{cY}}{\sigma_{aX}}\\
b1_{WV}&=\pm r_{XY}\frac{\sigma_{cY}}{\sigma_{aX}}\\
b1_{WV}&=\pm b1_{XY}\frac{c}{a}\\
\end{aligned}$$

$$\begin{aligned}
b0_{XY}&=\overline Y-b_1\overline X \\
b0_{WV}&=c\overline Y+d- (b1_{WV})(a\overline X-b)\\
\end{aligned}$$
    
###
##d.
```{r}
babySamp<-read.delim("/Users/Noah Cowit/Dropbox/Blue July 2016/College/Classes/S 3/STATS 346/babysamp.txt")

```

###i.)
```{r}
Y=babySamp$DBWT
X=babySamp$ESTGEST
lm(Y~X, data = babySamp)
```

###ii.)
```{r}
W=Y/454
V=X*7+14
lm(W~V, data = babySamp)
b1wv=((1/454)/7)*155.3
b1wv
b0wv=(1/454)*mean(Y)-b1wv*(mean(X)*7+14)
b0wv
```
The coefficients match.

##
   
#2
   
##
##a.
$$\begin{aligned}
E(b_1)&=E(\sum k_i y_i)\\
&=\sum k_i(E(\beta_0)+E(\beta_1x_i)+E(\epsilon))\\
&=\sum k_i(\beta_0+\beta_1x_i+0)\\
&=\sum k_i\beta_0+\sum k_i \beta_1 x_i\\
&=0+\sum \frac{x_i-\overline x}{S_{xx} } \beta_1 x_i\\
&=\frac {\beta_1}{S_{xx}}\sum (x_i-\overline x)x^i\\
&=\frac {S_{xx}\beta_1} {S_{xx}}\\
&=\beta_1
\end{aligned}$$
   
###
##b.
$$\begin{aligned}
Var(b_1)&=Var(\sum k_iy_i)\\
&=\sum k_i^2 Var(y_i)\\
&=\frac{S_{xx}}{S_{xx}^2}\sigma_\epsilon ^2\\
&=\frac{\sigma_{\varepsilon}^2}{S_{xx}}
\end{aligned}$$
  
##
    
#3
##a.
$$\begin{aligned}
Cov(y_i, b_i)&=Cov(y_i, \sum k_j y_j)\\
&=\sum k_j Cov(y_i,y_j)\\
&=k_i Cov(y_i,y_i)+\sum  k_j Cov(y_i,y_j)\ (where \ y_i \ and \ y_j \ are \ independent)\\
&=k_i Var(y_i)+0\\
&=k_i \sigma^2\\
\end{aligned}$$
    
###
##b.
$$\begin{aligned}
Cov(\overline y, b_i)&=Cov(\sum \frac {y_i}{N}, b_1)\\
&=\sum \frac{1}{N}k_i Cov(y_i,y_i)\\
&=\frac{1}{N}\sum k_i \sigma^2 \\
&=\frac{\sigma^2}{N}\sum k_i  \\
&=\frac{\sigma^2}{N}\sum \frac {x_i-\overline x}{S_{xx}} \\
&=\frac{\sigma^2}{N}*0\\
&=0\\
\end{aligned}$$
    
###
##c.
The slope of the regression would not change if you added 100 to every y value (which would increase $\overline y$ by 100). A change in mean has no effect on a change in slope.
    
###
##d.
$$\begin{aligned}
Cov(\overline y, b_i)&=Cov(\sum \frac {y_i}{N}, b_1)\\
&=Cov(b_0,b_1)\\
&=Cov(\overline Y-b_1 \overline x,b_1)\\
&=Cov(-b_1 \overline x,b_1)\\
&=-\overline xVar(b_1)\\
&=-\bar{x} \frac{\sigma^2}{S_{xx}}\\
\end{aligned}$$
    
###
##e.

$$\begin{aligned}
Var(b_0)&=Var(\bar Y-b)\\
&=Var(\bar Y)-\bar x^2 Var(b_1)\\
&=\frac {\sigma^2}{n}-x^2 \frac {\sigma^2}{S_{xx}}
\end{aligned}$$
    
###
##f.
$$\begin{aligned}
cor(b_0,b_1)&=\frac{ -\bar{x} \frac{\sigma^2}{S_{xx}}}{\sqrt{\frac {\sigma^2}{n}-x^2 \frac {\sigma^2}{S_{xx}}}\sqrt{ \frac {\sigma^2}{S_{xx}}}}\\
&=\frac{ -\bar{x} \frac{\sigma^2}{S_{xx}}}{\sqrt{\frac {1}{n {S_{xx}^2}}- \frac {x^2}{S_{xx}}}\sqrt{ {\sigma^4}}}\\
&=\frac{  \frac{-\bar{x}}{S_{xx}}}{\sqrt{\frac {1}{n {S_{xx}^2}}- \frac {x^2}{S_{xx}}}}*\frac{S_{xx}}{S_{xx}}\\
&= \frac{-\bar{x}}{\sqrt{\frac{S_{xx}}{n} + \bar{x}^2}}
\end{aligned}$$

##

#4
##a.
```{r}
bodyfat <- read.delim("http://sites.williams.edu/rdeveaux/files/2014/09/bodyfat.txt")
PBF<-bodyfat$Pct.BF
Weight<-bodyfat$Weight
lmfit<-lm(PBF~Weight,data=bodyfat)
```

###
##b.
```{r}
plot(Weight, PBF)
abline(lmfit)
```


###
##c.
```{r}
summary(lmfit)
qt(.975, df=248)
 0.18938-1.97*(.01533)
 0.18938+1.97*(.01533)
```
95% confidence interval for slope is (0.1591799,0.2195801).


##
##d.
```{r}
newdf <- data.frame(Weight=200)
predict(lmfit,new=newdf, se=T)
```

###
##e.
```{r}
plot(residuals(lmfit)~predict(lmfit))
```
There doesn't seem to be a relationship between the residuals and the predicted values.
##

#5
##a.
```{r}
library(mosaic)
slopes <- do(1000)*lm(Pct.BF~Weight,data=resample(bodyfat))
```
```{r}
hist<-hist(slopes$Weight)
hist
```
Yes, the center is at ~.19, very close to the .18938 we got on the regression earlier. The shape seems normal, which is to be expected in a bootstrap confidence interval. The distribution seems to hold true with the 95% confidence interval from earlier. Most of the slope values seem to be in the range of (0.1591799,0.2195801).

###
##b.
```{r}
quantile(slopes$Weight,.025)
quantile(slopes$Weight,.975)
```

###
##c.
The bootstrapped confidence interval (0.1602539,0.2181521) is very close to the CI from 4c (0.1591799,0.2195801).



